---
title: "02 Introduction to raretrans"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{02 Introduction to raretrans}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = FALSE, collapse = TRUE, comment = "#>")
library(raretrans)
```

## Overview

Matrix population models are a standard tool in ecology for projecting population
dynamics and quantifying the contributions of survival, growth, and reproduction to
population growth. A **population projection matrix** $\mathbf{A}$ is typically split
into two components:

- $\mathbf{T}$ — the **transition matrix**, whose columns describe the probabilities of
  surviving and moving between life-history stages (or staying in the same stage).
- $\mathbf{F}$ — the **fecundity matrix**, whose entries describe the average number of
  offspring contributed to each stage by each adult stage.

The full model is $\mathbf{A} = \mathbf{T} + \mathbf{F}$, and the asymptotic population
growth rate $\lambda$ is the dominant eigenvalue of $\mathbf{A}$.

### The structural-zero problem

When populations are small or study periods are short, some biologically possible
transitions are never observed — they appear as **zeros** in the data even though they
can and do occur in real populations. These are called **structural zeros** (or rare transitions). A
matrix containing structural zeros is mathematically problematic:

- It may fail to be **ergodic** (some stages become unreachable).
- It may be **reducible** (some subsets of stages are disconnected).
- The asymptotic growth rate $\lambda$ may be badly biased.

`raretrans` addresses this by using **Bayesian priors** to add small amounts of
"prior evidence" to the unobserved transitions, combining what we believe biologically
with what we actually observed.

---

## The example population: *Chamaedorea elegans*

Throughout this vignette we use a published matrix for the understorey palm
*Chamaedorea elegans* (parlour palm), compiled by Burns et al. (2013) and archived in
the COMPADRE Plant Matrix Database (Salguero-Gómez et al. 2015). The population has
three life-history stages:

| Stage label | Description |
|-------------|-------------|
| `seed`      | Propagule (seed) |
| `nonrep`    | Non-reproductive individual (seedling/juvenile) |
| `rep`       | Reproductive adult |

The transition ($\mathbf{T}$) and fecundity ($\mathbf{F}$) matrices are entered
directly below. This is the standard way to provide data to `raretrans` when you do
not have individual-level observations.

```{r define-matrices}
# ── Transition matrix T ────────────────────────────────────────────────────────
# Each column must sum to ≤ 1 (the remainder is mortality).
# Rows = destination stage; Columns = source stage.
T_mat <- matrix(
  c(0.2368, 0.0000, 0.0000,   # seed  → seed, nonrep, rep
    0.0010, 0.0100, 0.0200,   # nonrep → seed, nonrep, rep
    0.0000, 0.1429, 0.1364),  # rep  → seed, nonrep, rep
  nrow = 3, ncol = 3, byrow = TRUE,
  dimnames = list(c("seed", "nonrep", "rep"),
                  c("seed", "nonrep", "rep"))
)

# ── Fecundity matrix F ─────────────────────────────────────────────────────────
# F[i,j] = average number of stage-i individuals produced by one stage-j parent.
F_mat <- matrix(
  c(0, 0, 1.9638,   # reproductive adults produce seeds
    0, 0, 8.3850,   # reproductive adults produce nonreproductive offspring
    0, 0, 0.0000),  # no fecundity contributions to the reproductive stage
  nrow = 3, ncol = 3, byrow = TRUE,
  dimnames = list(c("seed", "nonrep", "rep"),
                  c("seed", "nonrep", "rep"))
)

# ── Wrap in a named list — the format expected by raretrans ────────────────────
TF <- list(T = T_mat, F = F_mat)
TF
```

Each **column** of $\mathbf{T}$ describes one source stage. For example, the column
`seed` tells us that in one time step a seed has:

- a 23.7% chance of remaining a seed (`seed` → `seed`),
- a 0.1% chance of germinating and becoming non-reproductive (`seed` → `nonrep`), and
- the remaining ~76% chance of dying (not shown; implied by the column sum < 1).

The $\mathbf{F}$ matrix shows that only reproductive adults (`rep`) contribute
offspring: each adult produces on average 1.96 seeds and 8.39 non-reproductive
individuals per time step. These large fecundity values are typical for tropical
palms that invest heavily in reproduction.

---

## Providing the stage-population vector *N*

Many functions in `raretrans` require the **starting number of individuals** observed
in each stage at the beginning of the census period. This vector $\mathbf{N}$ is used
to convert the transition probabilities back to raw counts (because the Bayesian model
works on counts, not proportions).

If you have individual-level data in a data frame you can use `get_state_vector()` to
extract $\mathbf{N}$ automatically (see the `onepopperiod` vignette for a worked
example). Here we supply it directly, using a hypothetical census of 80 seeds, 25
non-reproductive plants, and 12 reproductive adults.

```{r define-N}
N <- c(seed = 80, nonrep = 25, rep = 12)
N
```

---

## Part 1: Observed matrix and its limitations

Before adding any priors, let us examine the raw matrix.

```{r observed-lambda}
# Dominant eigenvalue = asymptotic growth rate λ
A_obs <- TF$T + TF$F
lambda_obs <- Re(eigen(A_obs)$values[1])
cat("Observed lambda:", round(lambda_obs, 3), "\n")
```

Now look at the column sums of $\mathbf{T}$:

```{r col-sums}
colSums(TF$T)
```

The `seed` column sums to only `r round(sum(TF$T[,1]), 4)`. The rest (`r round(1 - sum(TF$T[,1]), 4)`) represents mortality of seeds. Importantly, the transition from `seed` to `nonrep` is 0.001 — very small and easily missed in a short study. In a small sample we might observe zero such transitions, leaving a **structural zero** in the data matrix even though we know germination happens.

Similarly, the `nonrep` → `rep` entry (0.1429) and the back-transition `rep` → `nonrep` (0.0200) could be missed in small populations. When structural zeros appear in real data, we need a principled way to incorporate our biological knowledge that these transitions can occur.

---

## Part 2: Adding priors with `fill_transitions()`

### Uniform (non-informative) prior

The simplest prior is the **uniform Dirichlet** prior. It adds a small equal weight
to every possible fate. For a 3-stage system there are 4 fates per column
(3 stages + death), so a uniform prior with weight 1 adds 0.25 to each fate count.

The prior matrix `P` has one **more row than the transition matrix** $\mathbf{T}$: the
extra row represents death.

```{r uniform-prior}
# Uniform prior: equal weight to all fates (including death)
P_uniform <- matrix(0.25, nrow = 4, ncol = 3)
fill_transitions(TF, N, P = P_uniform)
```

Compare with the original $\mathbf{T}$:

```{r compare-uniform}
TF$T
```

The uniform prior has filled in all the zero entries, including the biologically
impossible `seed` → `rep` transition. **This is one weakness of a uniform prior: it
does not distinguish between rare-but-possible transitions and impossible ones**. This is an example and should **NOT** be used as a default prior in practice.

### Informative (expert) prior

A better approach is to use an **informative prior** that reflects biological
knowledge. The prior matrix still has 4 rows (3 stages + death), and ideally each
column sums to 1 so that the weight is easy to interpret.

For *C. elegans* we specify:

- Seeds have a high chance of dying (0.70), a moderate chance of staying as a seed
  (0.25), and a small chance of germinating to non-reproductive (0.05); the direct
  jump to reproductive is impossible (0.00).
- Non-reproductive plants mostly stay non-reproductive (0.55) or die (0.25); they can
  grow to reproductive (0.15) but very rarely revert to seed (0.05).
- Reproductive adults mostly stay reproductive (0.70) or die (0.10); they can revert
  to non-reproductive (0.15) or, very rarely, to seed (0.05).

```{r informative-prior}
P_info <- matrix(
  c(0.25, 0.05, 0.00,   # → seed
    0.05, 0.55, 0.15,   # → nonrep
    0.00, 0.15, 0.70,   # → rep
    0.70, 0.25, 0.15),  # → dead
  nrow = 4, ncol = 3, byrow = TRUE
)
# Verify columns sum to 1
colSums(P_info)
```

Now apply this prior with the default weight of 1 (equivalent to adding 1 individual's
worth of prior "observations"):

```{r informative-fill}
T_posterior <- fill_transitions(TF, N, P = P_info)
T_posterior
```

The informative prior fills in the rare transitions without creating the impossible
`seed` → `rep` entry (which stays at 0 because we set that prior entry to 0).

### Adjusting prior weight

The `priorweight` argument controls how strongly the prior influences the posterior.
A weight of 1 (the default) means the prior counts as 1 individual. Setting
`priorweight = 0.5` halves that influence:

```{r priorweight}
fill_transitions(TF, N, P = P_info, priorweight = 0.5)
```

With `priorweight = 0.5` and $N = 25$ non-reproductive plants observed, the effective
prior sample size for that column is $0.5 \times 25 = 12.5$ individuals, still less
than the data. This is the recommended approach: **the prior sample size should be
smaller than the observed sample size** so that the data dominate the posterior.

---

## Part 3: Adding priors for fecundity with `fill_fertility()`

The fecundity prior follows a **Gamma-Poisson** (negative binomial) model. We specify
two matrices, `alpha` and `beta`, whose entries are the shape and rate parameters of
the Gamma prior on the Poisson rate. Use `NA` for entries that represent stages that
cannot reproduce or fates that do not arise from reproduction.

```{r fertility-prior}
# alpha = prior "offspring counts"; beta = prior "adult counts"
# Only the (seed, rep) and (nonrep, rep) entries are reproductive
alpha <- matrix(
  c(NA, NA, 1e-5,
    NA, NA, 1e-5,
    NA, NA, NA),
  nrow = 3, ncol = 3, byrow = TRUE
)
beta <- matrix(
  c(NA, NA, 1e-5,
    NA, NA, 1e-5,
    NA, NA, NA),
  nrow = 3, ncol = 3, byrow = TRUE
)

F_posterior <- fill_fertility(TF, N, alpha = alpha, beta = beta)
F_posterior
```

The tiny values of `alpha` and `beta` (1e-5) produce a nearly non-informative prior:
the posterior fecundity is almost identical to the observed fecundity because the data
completely dominate. This is the correct behaviour when you have no prior information
about fecundity.

### Combined posterior matrix

```{r combined-posterior}
posterior <- list(
  T = fill_transitions(TF, N, P = P_info),
  F = fill_fertility(TF, N, alpha = alpha, beta = beta)
)
posterior

# Asymptotic growth rate of the posterior matrix
lambda_post <- Re(eigen(posterior$T + posterior$F)$values[1])
cat("Posterior lambda:", round(lambda_post, 3), "\n")
cat("Observed lambda:", round(lambda_obs, 3), "\n")
```

The posterior $\lambda$ is slightly different from the observed $\lambda$ because the
prior redistributes a small amount of probability mass from the observed transitions to
the rare ones.

---

## Part 4: Other return types

Both `fill_transitions()` and `fill_fertility()` accept a `returnType` argument for
accessing different representations of the posterior.

### Augmented fate matrix (TN)

Setting `returnType = "TN"` returns the **augmented fate matrix** — the raw Dirichlet
posterior counts (not divided by column sums). The extra bottom row gives the posterior
count for the death fate.

```{r returnType-TN}
TN <- fill_transitions(TF, N, P = P_info, returnType = "TN")
TN
```

This is useful for simulation (see Part 5) and for computing credible intervals
directly (see Part 6).

### Full projection matrix A

Setting `returnType = "A"` returns $\mathbf{T} + \mathbf{F}$ directly:

```{r returnType-A}
fill_transitions(TF, N, P = P_info, returnType = "A")
```

### Alpha and beta vectors (fertility)

Setting `returnType = "ab"` in `fill_fertility()` returns the posterior Gamma
parameters:

```{r returnType-ab}
fill_fertility(TF, N, alpha = alpha, beta = beta, returnType = "ab")
```

---

## Part 5: Simulating matrices and credible intervals on $\lambda$

Point estimates of $\lambda$ do not convey uncertainty. To obtain a **credible
interval** on the asymptotic growth rate we simulate many matrices from the posterior
distribution using `sim_transitions()`.

```{r sim-lambda, fig.width=6, fig.height=4, fig.cap="Posterior distribution of the asymptotic growth rate λ for *Chamaedorea elegans*. The dashed vertical line marks λ = 1 (stable population). Values to the right indicate growth; values to the left indicate decline."}
set.seed(20240301)  # for reproducibility

# Simulate 1000 projection matrices from the posterior
sims <- sim_transitions(TF, N,
                        P     = P_info,
                        alpha = alpha,
                        beta  = beta,
                        priorweight = 0.5,
                        samples = 1000)

# Extract λ from each simulated matrix
lambdas <- sapply(sims, function(mat) Re(eigen(mat)$values[1]))

# Summarise
cat("Posterior median λ:", round(median(lambdas), 3), "\n")
cat("95% credible interval: [",
    round(quantile(lambdas, 0.025), 3), ",",
    round(quantile(lambdas, 0.975), 3), "]\n")
cat("P(λ > 1):", round(mean(lambdas > 1), 3), "\n")

# Plot
hist(lambdas,
     breaks = 30,
     main   = expression("Posterior distribution of " * lambda),
     xlab   = expression(lambda),
     col    = "steelblue", border = "white")
abline(v = 1, lty = 2, lwd = 2, col = "firebrick")
```

The histogram shows the full posterior uncertainty about $\lambda$. `P(λ > 1)` is the
**posterior probability that the population is growing** — a more informative summary
than a point estimate alone.

---

## Part 6: Credible intervals on individual transition probabilities

### Computing credible intervals with `transition_CrI()`

`transition_CrI()` computes the marginal posterior **beta** credible intervals for
every entry in $\mathbf{T}$, including the probability of dying. It returns a tidy
data frame with one row per source–destination pair.

```{r cri}
cri <- transition_CrI(TF, N,
                      P           = P_info,
                      priorweight = 0.5,
                      ci          = 0.95,
                      stage_names = c("seed", "nonrep", "rep"))
cri
```

Each row gives the posterior mean transition probability and its 95% symmetric
credible interval. The `dead` rows summarise the probability of mortality from each
source stage.

### Visualising credible intervals with `plot_transition_CrI()`

```{r plot-cri, fig.width=7, fig.height=4, fig.cap="Posterior mean transition probabilities (points) and 95% credible intervals (lines) for *Chamaedorea elegans*. Each panel shows the fate distribution from one source stage."}
plot_transition_CrI(cri)
```

You can hide the death fate if you prefer to focus on the living stages:

```{r plot-cri-no-dead, fig.width=7, fig.height=4, fig.cap="As above but excluding the dead fate."}
plot_transition_CrI(cri, include_dead = FALSE)
```

Notice that the credible intervals for the `seed` → `nonrep` transition are very wide,
reflecting the fact that this germination event is rarely observed (only 0.1% per time
step in this matrix). The prior provides a small regularising signal but the data are
sparse, so uncertainty is high.

### Visualising the full posterior density with `plot_transition_density()`

`plot_transition_density()` arranges the full marginal posterior **beta density** for
every transition as a matrix of panels, mirroring the layout of $\mathbf{T}$. Each
column corresponds to a source stage and each row to a destination stage. The shaded
region is the 95% credible interval.

```{r plot-density, fig.width=7, fig.height=7, fig.cap="Posterior beta density for each transition in the *C. elegans* matrix. Columns = source stage (from); rows = destination stage (to). Shaded region = 95% credible interval. Zero-probability transitions show a degenerate spike at 0."}
plot_transition_density(TF, N,
                        P           = P_info,
                        priorweight = 0.5,
                        stage_names = c("seed", "nonrep", "rep"))
```

Panels on the main diagonal (e.g., `seed` → `seed`, `rep` → `rep`) tend to have
well-constrained distributions because stasis is commonly observed. Off-diagonal
transitions that are biologically rare show broad, flat densities — exactly where the
prior matters most.

Excluding the death row gives a tighter layout focused on the living stages:

```{r plot-density-no-dead, fig.width=7, fig.height=5, fig.cap="As above but excluding the dead fate row."}
plot_transition_density(TF, N,
                        P             = P_info,
                        priorweight   = 0.5,
                        stage_names   = c("seed", "nonrep", "rep"),
                        include_dead  = FALSE)
```

---

## Summary

This vignette introduced the core workflow of `raretrans`:

1. **Define** your transition matrix $\mathbf{T}$ and fecundity matrix $\mathbf{F}$,
   either from a data frame (using `popbio::projection.matrix()` and
   `get_state_vector()`) or by entering them directly.
2. **Specify a prior** matrix `P` that encodes biological knowledge about which
   transitions are possible and how likely they are. Prefer informative priors over the
   uniform default when you have relevant expertise.
3. **Fill structural zeros** using `fill_transitions()` and `fill_fertility()`.
4. **Simulate** many matrices from the posterior with `sim_transitions()` to obtain a
   credible interval on $\lambda$.
5. **Visualise** transition-level uncertainty with `transition_CrI()`,
   `plot_transition_CrI()`, and `plot_transition_density()`.

For a worked example that starts from individual-level field observations see
`vignette("onepopperiod")`.

---

## References

Burns, J., Blomberg, S., Crone, E., Ehrlén, J., Knight, T., Pichancourt, J.-B.,,
 Ramula, S., Wardle, G., & Buckley, Y. (2010). Empirical tests of life-history,
evolution theory using phylogenetic analysis of plant demography.,
 *Journal of Ecology*, **98**(2), 334–344.,
 <https://doi.org/10.1111/j.1365-2745.2009.01634.x>

Salguero-Gómez, R., Jones, O. R., Archer, C. R., Buckley, Y. M., Che-Castaldo, J.,
  Caswell, H., ..., & Vaupel, J. W. (2015). The COMPADRE Plant Matrix Database: an
  open online repository for plant demography. *Journal of Ecology*, **103**, 202–218.
  <https://doi.org/10.1111/1365-2745.12334>

Tremblay, R. L., Tyre, A. J., Pérez, M.-E., & Ackerman, J. D. (2021). Population projections from holey matrices: Using prior information to estimate rare transition events. Ecological Modelling, 447, 109526. https://doi.org/10.1016/j.ecolmodel.2021.109526
