---
title: "Simulation Studies"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulation Studies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(huxtable.add_colnames = FALSE,
        digits = 3)
```

```{r setup}
library(raretrans)
library(tidyverse)
library(popbio)
library(popdemo)
library(huxtable)
library(furrr)

# Raymond's theme modifications
rlt_theme <- theme(axis.title.y = element_text(colour="grey20",size=15,face="bold"),
        axis.text.x = element_text(colour="grey20",size=15, face="bold"),
        axis.text.y = element_text(colour="grey20",size=15,face="bold"),  
        axis.title.x = element_text(colour="grey20",size=15,face="bold"))
```

Reviewers have asked for a simulation study to confirm that the approach works. 
We will start with two different "true" matrices, simulate observed data from
these, and then use different priors to evaluate how the method works. 

# Calathea, an understory herb

The first example comes from `popbio::calathea` containing matrices published in
Horvitz and Schemske (1995). There are 4 plots and up to 5 years of transitions,
as well as a pooled matrix. Many of the transitions are very small probabilities. The pooled matrix is both ergodic and irreducible, and has a population growth rate of 
`r (true_lambda <- lambda(calathea$pooled))`. 


```{r}
calathea$pooled
```

We begin by simulating a data frame with N observations of plants. 
The initial population will be sampled from the stable stage distribution.
The function `raretrans::sim_transition` doesn't quite do what we 
want because it returns matrices, and we want the underlying observations.
Maybe that's what sim_transition should do, but it doesn't.

```{r}
TF <- splitA(calathea$pooled, r = 1, c = 3:8)
df <- sim_observations(TF)
tail(df) # more interesting
```

```{r}
projection_matrix(df, stage = "stage", fate = "fate", fertility = "Sd")
```
OK, so seems to work with the modified `projection_matrix`. Now we
generate a number of samples to work with, calculating some 
matrix statistics along the way. I'll ignore seeds, using the `add` argument to boot_transitions, and inserting a `(0.5, 0.1, 0.01, 0.39)` prior for the Bayesian version.




```{r, warning=FALSE}
add_Sd <- c(1,1, TF$T[1,1], 2, 1, TF$T[2,1], 3, 1, TF$T[3,1])
prior_Sd <- c(0.5, 0.1, 0.01, 0.39)
results <- tibble(observations = map(1:100, function(x)sim_observations(TF, N = 100, initial_stages = c(0, 1, 1, 1, 1, 1, 1, 1))),
       TF = map(observations, projection_matrix, fertility = "Sd", TF = TRUE),
       N = map(observations, get_state_vector),
       Tunif = map2(TF, N, fill_transitions),
       Funif = map2(TF, N, fill_fertility),
       A = map(TF, ~.x$T+.x$F),
       Aunif = map2(Tunif, Funif, ~.x + .y)) %>% 
  mutate(lambdaraw = map_dbl(A, lambda),
         lambdaunif = map_dbl(Aunif, lambda))
ggplot(data = results) + 
  geom_point(mapping = aes(x = lambdaraw, y = lambdaunif)) + 
  geom_hline(yintercept = lambda(TF$T + TF$F))
```
The uniform prior clearly causes issues by creating values of 1 in the fertility 
matrix for all stages that are not represented in the sample. 

Now we need to create some priors for this plant. Start with samples of size 1
for transitions known to be possible. 

fix it so prior is > 0 but very small unless zeros are "structural".
Structural zeros should never be updated even if an observation is made. 

Fertility priors will be 0 except 
for stages known to reproduce, and these will be set close to zero.

```{r}
set.seed(9347905)
true_lambda <- lambda(TF$T + TF$F)
add_Sd <- c(1,1, TF$T[1,1], 2, 1, TF$T[2,1], 3, 1, TF$T[3,1])
transition_prior1 <- rbind(TF$T > 0, rep(1, times = ncol(TF$T)))
column_sums <- apply(transition_prior1, 2, sum)
transition_prior1 <- sweep(transition_prior1, 2, column_sums, `/` )
fertility_alpha <- TF$F * 0
fertility_alpha[1, 3:8] <- 1e-5
fertility_beta <- TF$F * 0 + (1-1e-5) # to keep sum = 1 
results <- tibble(observations = map(1:100, function(x)sim_observations(TF, N = 100, initial_stages = c(0, 1, 1, 1, 1, 1, 1, 1))),
       TF = map(observations, projection_matrix, fertility = "Sd", TF = TRUE, add = add_Sd),
       N = map(observations, get_state_vector),
       Tunif = map2(TF, N, fill_transitions, P = transition_prior1),
       Funif = map2(TF, N, fill_fertility, alpha = fertility_alpha, beta = fertility_beta),
       A = map(TF, ~.x$T+.x$F),
       Aunif = map2(Tunif, Funif, ~.x + .y)) %>% 
  mutate(lambdaraw = map_dbl(A, lambda),
         bootraw = map(observations, boot_transitions, iterations = 1000, add = add_Sd),
         lci_raw = map_dbl(bootraw, ~quantile(.$lambda, p = 0.025)),
         uci_raw = map_dbl(bootraw, ~quantile(.$lambda, p = 0.975)),
         lambdaunif = map_dbl(Aunif, lambda))
results <- mutate(results,
                  goodci = lci_raw < true_lambda & uci_raw > true_lambda,
                  sensibleci = lci_raw < lambdaraw & uci_raw > lambdaraw)


ggplot(data = results) +
  geom_point(mapping = aes(x = 1:100, y = lambdaraw, color = goodci)) + 
  geom_errorbar(mapping = aes(x = 1:100, ymin = lci_raw, ymax = uci_raw, color = goodci))+
  geom_hline(yintercept = lambda(TF$T + TF$F), linetype = 2, color = "blue")
```
So in many cases the observed lambda is on the edge of the interval, but at 
they are all at least "inside" now that I'm using the correct version of 
`projection_matrix` in the bootstrap.

What about a larger sample size and assuming knowledge of the seed stage transitions?

```{r}
set.seed(384768697)
initial_stages <- c(0, 1, 1, 1, 1, 1, 1, 1)
true_lambda <- lambda(TF$T + TF$F)
results <- tibble(observations = map(1:100, 
                                     function(x) sim_observations(TF, N = 500, initial_stages = initial_stages)),
       TF = map(observations, projection_matrix, fertility = "Sd", TF = TRUE, add = add_Sd),
       N = map(observations, get_state_vector),
       Tunif = map2(TF, N, fill_transitions, P = transition_prior1),
       Funif = map2(TF, N, fill_fertility, alpha = fertility_alpha, beta = fertility_beta),
       A = map(TF, ~.x$T+.x$F),
       Aunif = map2(Tunif, Funif, ~.x + .y)) %>% 
  mutate(lambdaraw = map_dbl(A, lambda),
         bootraw = map(observations, boot_transitions, iterations = 1000, add = add_Sd),
         lci_raw = map_dbl(bootraw, ~quantile(.$lambda, p = 0.025)),
         uci_raw = map_dbl(bootraw, ~quantile(.$lambda, p = 0.975)),
         lambdaunif = map_dbl(Aunif, lambda))
results <- mutate(results,
                  goodci = lci_raw < true_lambda & uci_raw > true_lambda,
                  sensibleci = lci_raw < lambdaraw & uci_raw > lambdaraw)
ggplot(data = results) +
  geom_point(mapping = aes(x = 1:100, y = lambdaraw)) + 
  geom_errorbar(mapping = aes(x = 1:100, ymin = lci_raw, ymax = uci_raw,
                              color = goodci)) +
  geom_hline(yintercept = lambda(TF$T + TF$F), linetype = 2, color = "blue")+
  scale_color_manual(values=c("FALSE" = "#F8766D", "TRUE" = "#00BA38"))
```

With a sample of 500 individuals `r sum(results$goodci)` intervals
include the true population growth rate. Indeed, only `r sum(results$sensibleci)`
intervals even include the observed growth rate. 

## Intervals from Bayesian posterior

```{r}
results <- results %>% 
  mutate(sim_TF = map2(TF, N, function(.x, .y) sim_transitions(TF = .x, 
                      N = .y, P = transition_prior1, 
                      alpha = fertility_alpha, beta = fertility_beta, 
                      samples = 1000)))
results <- results %>% 
  mutate(sim_lambdas = map(sim_TF, function(ll) map_dbl(ll, lambda)),
         sim_lci = map_dbl(sim_lambdas, quantile, p = 0.025),
         sim_uci = map_dbl(sim_lambdas, quantile, p = 0.975),
         good_sci = sim_lci < true_lambda & sim_uci > true_lambda,
         sensible_sci = sim_lci < lambdaunif & sim_uci > lambdaunif)
ggplot(data = results) +
  geom_point(mapping = aes(x = 1:100, y = lambdaunif)) + 
  geom_errorbar(mapping = aes(x = 1:100, ymin = sim_lci, ymax = sim_uci,
                              color = good_sci)) +
  geom_hline(yintercept = lambda(TF$T + TF$F), linetype = 2, color = "blue")+
  scale_color_manual(values=c("FALSE" = "#F8766D", "TRUE" = "#00BA38"))
```

The coverage property is terrible, and the estimated lambdas are very biased. However, we have not used all the information that was available to the bootstrap -- the prior for seedling transitions was uniform over the possible transitions. What if we switch to a more 
informative prior on seedling transitions? 

```{r}
transition_prior2 <- transition_prior1
prior_Sd <- c(0.5, 0.1, 0.01, 0.39)
transition_prior2[c(1:3, 9), 1] <- prior_Sd
results2 <- results %>% 
  mutate(sim_TF = map2(TF, N, function(.x, .y) sim_transitions(TF = .x, 
                      N = .y, P = transition_prior2, 
                      alpha = fertility_alpha, beta = fertility_beta, 
                      samples = 1000)))
results <- results %>% 
  mutate(sim_lambdas = map(sim_TF, function(ll) map_dbl(ll, lambda)),
         sim_lci = map_dbl(sim_lambdas, quantile, p = 0.025),
         sim_uci = map_dbl(sim_lambdas, quantile, p = 0.975),
         good_sci = sim_lci < true_lambda & sim_uci > true_lambda,
         sensible_sci = sim_lci < lambdaunif & sim_uci > lambdaunif)
ggplot(data = results) +
  geom_point(mapping = aes(x = 1:100, y = lambdaunif)) + 
  geom_errorbar(mapping = aes(x = 1:100, ymin = sim_lci, ymax = sim_uci,
                              color = good_sci)) +
  geom_hline(yintercept = lambda(TF$T + TF$F), linetype = 2, color = "blue")+
  scale_color_manual(values=c("FALSE" = "#F8766D", "TRUE" = "#00BA38"))
```

What about relative width and bias of the intervals versus sample size?

# relative width and bias as a function of sample size

```{r}

data(method_comparison_1)

# transition_prior3 <- transition_prior1
# prior_Sd <- c(0.5, 0.1, 0.01, 0.39)
# transition_prior3[c(1:3, 9), 1] <- prior_Sd * 1000
# 
# set.seed(817516)
# bigresults <- tibble(sample_size = seq(50, 500, by = 50),
#                      observations = map(sample_size, ~map(1:100,
#                                      function(x) sim_observations(TF, N = .x, initial_stages = initial_stages)))) %>%
#   unnest(cols = observations) %>%
#   mutate(
#        TF = map(observations, projection_matrix, fertility = "Sd", TF = TRUE, add = add_Sd),
#        N = map(observations, get_state_vector),
#        A = map(TF, ~.x$T+.x$F),
#        lambdaraw = map_dbl(A, lambda))
# 
# method_comparison_1 <- bigresults %>%
#   mutate(boot_TF = map(observations, boot_transitions, iterations = 1000, add = add_Sd),
#          sim_TF = map2(TF, N, function(.x, .y) sim_transitions(TF = .x,
#                       N = .y, P = transition_prior3,
#                       alpha = fertility_alpha, beta = fertility_beta,
#                       samples = 1000)))

bigresults2 <- method_comparison_1 %>% 
  mutate(boot_lambdas = map(boot_TF, "lambda"),
         boot_lci = map_dbl(boot_TF, ~quantile(.$lambda, p = 0.025)),
         boot_uci = map_dbl(boot_TF, ~quantile(.$lambda, p = 0.975)),
         good_bci = boot_lci < true_lambda & boot_uci > true_lambda,
         width_bci = (boot_uci - boot_lci) / lambdaraw,
         mean_bias_bci = map_dbl(boot_lambdas, ~mean(.x - true_lambda)),
         se_bias_bci = map_dbl(boot_lambdas, ~sd(.x - true_lambda)),
         sim_lambdas = map(sim_TF, function(ll) map_dbl(ll, lambda)),
         sim_lci = map_dbl(sim_lambdas, quantile, p = 0.025),
         sim_uci = map_dbl(sim_lambdas, quantile, p = 0.975),
         good_sci = sim_lci < true_lambda & sim_uci > true_lambda,
         width_sci = (sim_uci - sim_lci) / lambdaraw,
         mean_bias_sci = map_dbl(sim_lambdas, ~mean(.x - true_lambda)),
         se_bias_sci = map_dbl(sim_lambdas, ~sd(.x - true_lambda)))

figure_results <- bigresults2 %>% 
  group_by(sample_size) %>% 
  summarize(cover_boot = sum(good_bci),
            width_boot = mean(width_bci),
            width.lci_boot = quantile(width_bci, p = 0.025), 
            width.uci_boot = quantile(width_bci, p = 0.975), 
            bias_boot = mean(mean_bias_bci/true_lambda),
            bias.lci_boot = quantile(mean_bias_bci/true_lambda, p = 0.025), 
            bias.uci_boot = quantile(mean_bias_bci/true_lambda, p = 0.975), 
            cover_sim = sum(good_sci),
            width_sim = mean(width_sci),
            width.lci_sim = quantile(width_sci, p = 0.025), 
            width.uci_sim = quantile(width_sci, p = 0.975), 
            bias_sim = mean(mean_bias_sci),
            bias.lci_sim = quantile(mean_bias_sci/true_lambda, p = 0.025), 
            bias.uci_sim = quantile(mean_bias_sci/true_lambda, p = 0.975)) %>% 
  pivot_longer(cols = -1, 
              names_to = c(".value", "method"),
              names_sep = "_")

egg::ggarrange(
  ggplot(figure_results) + 
  geom_point(mapping = aes(x = sample_size, y = width, color = method)) + 
  geom_line(mapping = aes(x = sample_size, y = width, color = method)) + 
  geom_ribbon(mapping = aes(x = sample_size, ymin = width.lci, ymax = width.uci, fill = method), alpha = 0.1),
  ggplot(figure_results) + 
  geom_point(mapping = aes(x = sample_size, y = bias, color = method)) + 
  geom_line(mapping = aes(x = sample_size, y = bias, color = method)) + 
  geom_ribbon(mapping = aes(x = sample_size, ymin = bias.lci, ymax = bias.uci, fill = method), alpha = 0.1))
```

So when we provide the Bayesian method with a strong and accurate prior on the seedling transitions, as is assumed with the bootstrap, the Bayesian method
performs slightly better at small sample sizes than the non-parametric bootstrap. This means that the bootstrap with the assumed transitions is overly optimistic 
about the precision, and potentially biased.

# what about conflicting priors?

One thing that might concern users is if the prior knowledge is incorrect or
misleading in some way. What are some ways this might happen? Missing possible
transitions? Including impossible transitions? We rerun the simulations above
with two different possible errors in the prior information. First, we will 
use the weak prior on seedling transitions. In the second scenario we set to zero all 
prior transitions with probabilities smaller than 0.01, and perturb 
all other positive matrix entries by up to 25 or 50%. We compare both of these against the Bayesian method with accurate but uninformative priors on most transitions,
and accurate and strong priors on seedling transitions (closest to the bootstrap).

```{r}

#data(method_comparison_2)

transition_prior4 <- TF$T 
transition_prior4[TF$T < 0.01] <- 0
transition_prior4 <- transition_prior4 * 1.25 # 25% increase on priors
column_sums <- apply(transition_prior4, 2, sum)
transition_prior4[,column_sums > 1] <- sweep(transition_prior4[,column_sums > 1], 2, 
                                             column_sums[column_sums > 1], `/`)
column_sums <- apply(transition_prior4, 2, sum)
transition_prior4 <- rbind(transition_prior4, 1 - column_sums)


set.seed(449535)
plan("multisession", workers = 4)
bigresults <- tibble(sample_size = seq(50, 500, by = 50),
                     observations = future_map(sample_size, ~map(1:100,
                                     function(x) sim_observations(TF, N = .x,
                                                                  initial_stages = initial_stages)))) %>%
  unnest(cols = observations) %>%
  mutate(
       TF = map(observations, projection_matrix, fertility = "Sd", TF = TRUE, add = add_Sd),
       N = map(observations, get_state_vector),
       A = map(TF, ~.x$T+.x$F),
       lambdaraw = map_dbl(A, lambda))

method_comparison_2 <- bigresults %>%
  mutate(sim2_TF = map2(TF, N, function(.x, .y) sim_transitions(TF = .x,
                      N = .y, P = transition_prior1,
                      alpha = fertility_alpha, beta = fertility_beta,
                      samples = 1000)),
         sim3_TF = map2(TF, N, function(.x, .y) sim_transitions(TF = .x,
                      N = .y, P = transition_prior4,
                      alpha = fertility_alpha, beta = fertility_beta,
                      samples = 1000)))

bigresults2 <- method_comparison_2 %>% 
  mutate(boot_lambdas = map(boot_TF, "lambda"),
         boot_lci = map_dbl(boot_TF, ~quantile(.$lambda, p = 0.025)),
         boot_uci = map_dbl(boot_TF, ~quantile(.$lambda, p = 0.975)),
         good_bci = boot_lci < true_lambda & boot_uci > true_lambda,
         width_bci = (boot_uci - boot_lci) / lambdaraw,
         mean_bias_bci = map_dbl(boot_lambdas, ~mean(.x - true_lambda)),
         se_bias_bci = map_dbl(boot_lambdas, ~sd(.x - true_lambda)),
         sim_lambdas = map(sim_TF, function(ll) map_dbl(ll, lambda)),
         sim_lci = map_dbl(sim_lambdas, quantile, p = 0.025),
         sim_uci = map_dbl(sim_lambdas, quantile, p = 0.975),
         good_sci = sim_lci < true_lambda & sim_uci > true_lambda,
         width_sci = (sim_uci - sim_lci) / lambdaraw,
         mean_bias_sci = map_dbl(sim_lambdas, ~mean(.x - true_lambda)),
         se_bias_sci = map_dbl(sim_lambdas, ~sd(.x - true_lambda)))

figure_results <- bigresults2 %>% 
  group_by(sample_size) %>% 
  summarize(cover_boot = sum(good_bci),
            width_boot = mean(width_bci),
            width.lci_boot = quantile(width_bci, p = 0.025), 
            width.uci_boot = quantile(width_bci, p = 0.975), 
            bias_boot = mean(mean_bias_bci/true_lambda),
            bias.lci_boot = quantile(mean_bias_bci/true_lambda, p = 0.025), 
            bias.uci_boot = quantile(mean_bias_bci/true_lambda, p = 0.975), 
            cover_sim = sum(good_sci),
            width_sim = mean(width_sci),
            width.lci_sim = quantile(width_sci, p = 0.025), 
            width.uci_sim = quantile(width_sci, p = 0.975), 
            bias_sim = mean(mean_bias_sci),
            bias.lci_sim = quantile(mean_bias_sci/true_lambda, p = 0.025), 
            bias.uci_sim = quantile(mean_bias_sci/true_lambda, p = 0.975)) %>% 
  pivot_longer(cols = -1, 
              names_to = c(".value", "method"),
              names_sep = "_")

egg::ggarrange(
  ggplot(figure_results) + 
  geom_point(mapping = aes(x = sample_size, y = width, color = method)) + 
  geom_line(mapping = aes(x = sample_size, y = width, color = method)) + 
  geom_ribbon(mapping = aes(x = sample_size, ymin = width.lci, ymax = width.uci, fill = method), alpha = 0.1),
  ggplot(figure_results) + 
  geom_point(mapping = aes(x = sample_size, y = bias, color = method)) + 
  geom_line(mapping = aes(x = sample_size, y = bias, color = method)) + 
  geom_ribbon(mapping = aes(x = sample_size, ymin = bias.lci, ymax = bias.uci, fill = method), alpha = 0.1))
```



# What about a different sampling strategy? 

If an ecologist uses a fixed area to get the sample, then using the stable stage
distribution would be a good initial vector. However, rarer stages could be
oversampled by searching for them. 

```{r}

set.seed(5786576)
initial_stages <- c(0.5, 1, 1, 1, 1, 2, 2, 5)
true_lambda <- lambda(TF$T + TF$F)
results <- tibble(observations = map(1:100, 
                                     function(x) sim_observations(TF, N = 500, initial_stages = initial_stages)),
       TF = map(observations, projection_matrix, fertility = "Sd", TF = TRUE),
       N = map(observations, get_state_vector),
       Tunif = map2(TF, N, fill_transitions, P = transition_prior),
       Funif = map2(TF, N, fill_fertility, alpha = fertility_alpha, beta = fertility_beta),
       A = map(TF, ~.x$T+.x$F),
       Aunif = map2(Tunif, Funif, ~.x + .y)) %>% 
  mutate(lambdaraw = map_dbl(A, lambda),
         bootraw = map(observations, boot_transitions, iterations = 1000),
         lci_raw = map_dbl(bootraw, ~quantile(.$lambda, p = 0.025)),
         uci_raw = map_dbl(bootraw, ~quantile(.$lambda, p = 0.975)),
         lambdaunif = map_dbl(Aunif, lambda))
results <- mutate(results,
                  goodci = lci_raw < true_lambda & uci_raw > true_lambda,
                  sensibleci = lci_raw < lambdaraw & uci_raw > lambdaraw)
ggplot(data = results) +
  geom_point(mapping = aes(x = 1:100, y = lambdaraw)) + 
  geom_errorbar(mapping = aes(x = 1:100, ymin = lci_raw, ymax = uci_raw,
                              color = goodci)) +
  geom_hline(yintercept = lambda(TF$T + TF$F), linetype = 2, color = "blue")
```

 How about the Bayesian approach with
oversampled data?


```{r}
# no set seed because using results from previous chunk.
results <- results %>% 
  mutate(sim_TF = map2(TF, N, function(.x, .y) sim_transitions(TF = .x, 
                      N = .y, P = transition_prior, 
                      alpha = fertility_alpha, beta = fertility_beta, 
                      samples = 1000)))
results <- results %>% 
  mutate(sim_lambdas = map(sim_TF, function(ll) map_dbl(ll, lambda)),
         sim_lci = map_dbl(sim_lambdas, quantile, p = 0.025),
         sim_uci = map_dbl(sim_lambdas, quantile, p = 0.975),
         good_sci = sim_lci < true_lambda & sim_uci > true_lambda,
         sensible_sci = sim_lci < lambdaunif & sim_uci > lambdaunif)
ggplot(data = results) +
  geom_point(mapping = aes(x = 1:100, y = lambdaunif)) + 
  geom_errorbar(mapping = aes(x = 1:100, ymin = sim_lci, ymax = sim_uci,
                              color = good_sci)) +
  geom_hline(yintercept = lambda(TF$T + TF$F), linetype = 2, color = "blue")
```



# Desert Tortoise

Doak et al. (1994) published four matrices for desert tortoise (*Gopherus agassizii*). 
These matrices have entries with relatively large values for all entries, except for the transition from adult1 to adult2, which is ~0.02.
We will try the bootstrap CI on these matrices.

```{r}
TF <- splitA(tortoise[["high"]])
set.seed(837563)
transition_prior <- rbind(TF$T > 0, rep(1, times = ncol(TF$T)))
column_sums <- apply(transition_prior, 2, sum)
transition_prior <- sweep(transition_prior, 2, column_sums, `/` )
fertility_alpha <- TF$F * 0
fertility_alpha[1, 6:8] <- 1e-5
fertility_beta <- TF$F * 0 + (1-1e-5) # to keep sum = 1 
results <- tibble(observations = map(1:100, function(x)sim_observations(TF, N = 500)),
       TF = map(observations, projection_matrix, fertility = "yearling", TF = TRUE),
       N = map(observations, get_state_vector),
       Tunif = map2(TF, N, fill_transitions, P = transition_prior),
       Funif = map2(TF, N, fill_fertility, alpha = fertility_alpha, beta = fertility_beta),
       A = map(TF, ~.x$T+.x$F),
       Aunif = map2(Tunif, Funif, ~.x + .y)) %>% 
  mutate(lambdaraw = map_dbl(A, lambda),
         bootraw = map(observations, boot_transitions, iterations = 1000, by.stage.counts = TRUE, fertility = "yearling"),
         lci_raw = map_dbl(bootraw, ~quantile(.$lambda, p = 0.025)),
         uci_raw = map_dbl(bootraw, ~quantile(.$lambda, p = 0.975)),
         lambdaunif = map_dbl(Aunif, lambda))

results$A
stable.stage(tortoise[["high"]])
all.sims <- bind_rows(results$observations)

ggplot(data = results) + 
  geom_point(mapping = aes(x = lambdaraw, y = lambdaunif)) + 
  geom_hline(yintercept = lambda(TF$T + TF$F)) + 
  geom_errorbarh(mapping = aes(xmin = lci_raw, xmax = uci_raw, y = lambdaunif))
```


# Caveats

Simulating matrices from the posterior ignores correlations between vital rates,
which might cause the credible intervals to be too narrow. I wonder if 
applying bootstrapped samples to the priors would be better? 
But the more I think about it, the less worried I am about that. I think the 
correlation between vital rates is in variation over time, and we’re just looking 
at a single transition. 




# Literature cited

Horvitz, C.C. and D.W. Schemske. 1995. Spatiotemporal variation in demographic transitions of a tropical understory herb: Projection matrix analysis. Ecological Monographs 65:155-192.
